### 2014-11-11 model.tex
edits

Figure~\ref{fig:sysmod} provides a schematic diagram of our system model. The
distributed store is composed of a collection of \emph{replicas}, each of which
stores a set of \emph{objects} ($x,y,\ldots$). We assume that every object is
replicated at every replica in the store. The state of an object at any replica
is the set of all updates (\emph{effects}) performed on the object. For
example, the state of $x$ at replica 1 is the set composed of effects $w^x_1$
and $w^x_2$.

@@

When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
replica 1, it simply \emph{reduces} over the current set of effects at that
replica on that object ($w^x_1$ and $w^x_2$), produces a result $v1$ that is
sent back to the client, and emits a \emph{single} new effect $w^x_4$ that is
appended to the state of $x$ at replica 1. Thus, every operation is evaluated
over a \emph{snapshot} of the state of the object on which it is invoked. In
this case, the effects $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
$\visZ$ is the visibility relation between effects. Visibility is an
irreflexive and asymmetric relation, and only relates effects produced by
operations on the same object. Executing a read-only operation is similar
except that no new effects are produced.

The effect added to a particular replica is asynchronously sent to other
replicas, and eventually merged into all other replicas. Two effects $w^x_4$
and $w^x_5$ that arise from the same session are said to be in \emph{session
order} (written logically as $\so{w^x_4}{w^x_5}$). Session order is an
irreflexive, transitive relation. The effects $w^x_4$ and $w^x_5$ arising from
operations applied to the same object $x$ are said to be under the \emph{same
object} relation, written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate
every effect with the operation that generated the effect with the help of a
relation $\operZ$. In the current example, $\oper{w^x_4}{foo}$ and
$\oper{w^x_5}{bar}$ hold. For simplicity, we assume all operation names across
all object types are distinct.

This model admits all the inconsistencies associated with eventual consistency.
The goal of this work is to identify the precise consistency level for each
operation such that application-level constraints are not violated. In the next
section, we will concretely describe the challenges associated with
constructing a consistent bank account on top of an eventually consistent data
store. Subsequently, we will illustrate how our contract and specification
language, armed with the primitive relations $\visZ$, $\soZ$, $\sameobjZ$ and
$\operZ$, mitigates these challenges.

>>>>

In this section, we describe the system model and introduce the primitive
relations that our contract language is seeded with. Figure~\ref{fig:sysmod}
provides a schematic diagram of our system model. The distributed store is
composed of a collection of \emph{replicas}, each of which stores a set of
\emph{objects} ($x,y,\ldots$). We assume that every object is replicated at
every replica in the store. The state of an object at any replica is the set of
all updates (\emph{effects}) performed on the object. For example, the state of
$x$ at replica 1 is the ordered set composed of effects $w^x_1$ and $w^x_2$.

@@

When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
replica 1, it simply \emph{reduces} over the current set of effects at that
replica on that object ($w^x_1$ and $w^x_2$), produces a result $v1$ that is
sent back to the client, and emits a \emph{single} new effect $w^x_4$ that is
appended to the state of $x$ at replica 1. Thus, every operation is evaluated
over a \emph{snapshot} of the state of the object on which it is invoked. In
this case, the effects $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
written logically as $\small \vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
$\small \visZ$ is the visibility relation between effects. Visibility is an
irreflexive and asymmetric relation, and only relates effects produced by
operations on the same object. Executing a read-only operation is similar
except that no new effects are produced.

The effect added to a particular replica is asynchronously sent to other
replicas, and eventually merged into all other replicas. Two effects $w^x_4$
and $w^x_5$ that arise from the same session are said to be in \emph{session
order} (written logically as $\small \so{w^x_4}{w^x_5}$). Session order is an
irreflexive, transitive relation. The effects $w^x_4$ and $w^x_5$ arising from
operations applied to the same object $x$ are said to be under the \emph{same
object} relation, written $\small \sameobj{w^x_4}{w^x_5}$. Finally, we can
associate every effect with the operation that generated the effect with the
help of a relation $\small \operZ$. In the current example, $\small
\oper{w^x_4}{foo}$ and $\small \oper{w^x_5}{bar}$ hold. For simplicity, we
assume all operation names across all object are distinct.

This model admits all the inconsistencies associated with eventual consistency.
The goal of this work is to identify the precise consistency level for each
operation such that application-level constraints are not violated. In the next
section, we will concretely describe the challenges associated with
constructing a consistent bank account on top of an eventually consistent data
store. Subsequently, we will illustrate how our contract and specification
language, armed with the primitive relations $\small \visZ$, $\small \soZ$,
$\small \sameobjZ$ and $\small \operZ$, mitigates these challenges.

### 2014-11-11 motivation.tex
edits

A key novelty in \name is that it allows the addition of new RDTs to the store,
which obviates the need for coercing the application logic to utilize the store
provided data types. In addition, \name treats the convergence semantics of the
data type separately from its consistency properties. This separation of
concerns permits \emph{operational} reasoning for conflict resolution, and
\emph{declarative} reasoning for consistency. The combination of these
techniques enhances the programmability of the store.

@@

@@

\begin{figure}
\centering
\subfigure[Unsafe withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
\hfill
\subfigure[Negative balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
\hfill
\subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
\caption{Anomalies possible under eventual consistency for the get balance operation.}
\label{fig:ba_anomalies}
\end{figure}

Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
that all operations in the figure are on the same bank account object with the
initial balance being zero. Session 1 performs a \cf{deposit} of 100, followed
by a \cf{withdraw} of 80 in the same session. The \cf{withdraw} operation
witnesses the deposit and succeeds\footnote{Although visibility and session
order relations relate effects, we have abused the notation in these examples
to relate operations, with the idea that the relations relate the effect
emitted by those operations}. Subsequently, session 2 perform a \cf{withdraw}
operation, but importantly, due to eventual consistency, only witnesses the
\cf{deposit} from session 1, but not the subsequent withdraw. Hence, this
\cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
constraint. A subsequent \cf{getBalance} operation, that happens to witness all
the previous operations, would report a negative balance.

@@

\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$
is an effect that was emitted either by a \cf{deposit} or a \cf{withdraw}
operation. If a \cf{withdraw} $b$ is visible to \cf{getBalance} $\cureff$, then
all \cf{deposit} operations $a$ visible to $b$ should also be visible to
$\cureff$. This prevents negative balance anomalies. Our contract language
provides operators to compose relations. The syntax $(R_1 \cap R_2)(a,b)$ is
equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the above contract
says that if a \cf{deposit} or a \cf{withdraw} operation precedes a
\cf{getBalance} operation in session order, and is applied on the same object
as the \cf{getBalance} operation, then it must be the case that the
\cf{getBalance} operation witnesses the effects of the preceding operations.

>>>>

A key novelty in \name is that it allows the addition of new RDTs to the store,
which obviates the need for coercing the application logic to utilize the store
provided data types. In addition, \name treats the convergence semantics (i.e.,
\emph{how} conflicting updates are resolved) of the data type separately from
its consistency properties (i.e., \emph{when} updates become visible). This
separation of concerns permits \emph{operational} reasoning for conflict
resolution, and \emph{declarative} reasoning for consistency. The combination
of these techniques enhances the programmability of the store.

@@

\subsubsection{Summarization}
\label{sec:summarize}

Observe that the definition of \cf{getBalance} reduces over the \emph{entire
history} of updates to an account. If we are to realize an efficient
implementation of this bank account RDT, we need a \emph{summary} of the
account history. Intuitively, the current account balance summarizes the state
of an account. A bank account with the history \cf{[Deposit 10, Withdraw 5]} is
\emph{observably equivalent} to a bank account with a single deposit operation
\cf{[Deposit 5]}; we can replace the earlier history with the latter and a
client of the store would not able to tell the difference between the two.

This notion of observable equivalence can be generalized to other RDTs as well.
For example, a last-writer-wins register with multiple updates is equivalent to
a register with only the last write. Similarly, a set with a collection of add
and remove operations is equivalent to a set with a series of additions of live
elements from the original set. Since the notion of observable equivalence is
specific to each RDT, we expect the programmer to provide a summarization
function (\cf{summarize}) of type \cf{[e] -> [e]} for each RDT. We utilize this
function to ensure that the state of an object remains summarized in each
replica. The summarization function for the bank account is {\footnotesize
\cf{summarize ctxt = [Deposit \$ sel1 \$ getBalance ctxt ()]}}.

@@

\begin{figure}[ht]
\centering
\subfigure[Unsafe withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
\hfill
\subfigure[Negative balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
\hfill
\subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
\caption{Anomalies possible under eventual consistency for the get balance operation.}
\label{fig:ba_anomalies}
\end{figure}

Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
that all operations in the figure are on the same bank account object with the
initial balance being zero. Session 1 performs a \cf{deposit} of 100, followed
by a \cf{withdraw} of 80 in the same session. The \cf{withdraw} operation
witnesses\footnote{Although visibility and session
order relations relate effects, we have abused the notation in these examples
to relate operations, with the idea that the relations relate the effect
emitted by those operations} the deposit and succeeds. Subsequently, session 2 perform a \cf{withdraw}
operation, but importantly, due to eventual consistency, only witnesses the
\cf{deposit} from session 1, but not the subsequent withdraw. Hence, this
\cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
constraint. A subsequent \cf{getBalance} operation, that happens to witness all
the previous operations, would report a negative balance.

@@

\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$
is an effect that was emitted either by a \cf{deposit} or a \cf{withdraw}
operation. If a \cf{withdraw} $b$ is visible to \cf{getBalance} $\cureff$, then
all \cf{deposit} operations $a$ visible to $b$ should also be visible to
$\cureff$. Intuitively, this prevents negative balance anomalies. Our contract
language provides operators to compose relations. The syntax $(R_1 \cap
R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
above contract says that if a \cf{deposit} or a \cf{withdraw} operation
precedes a \cf{getBalance} operation in session order, and is applied on the
same object as the \cf{getBalance} operation, then it must be the case that the
\cf{getBalance} operation witnesses the effects of the preceding operations.
